ODGI: understanding pangenome graphs

### 2021-11-03 manuscript/main.tex
reduce the warnings and the citation error (I'm skipping whatever remains of this entry)

\begin{itemize}
\item \textit{odgi build} constructs the ODGI data model from GFA file (\S\ref{sec:build}).
\item \textit{odgi view} converts the ODGI data model into GFA file (\S\ref{sec:build}).
\item \textit{odgi viz} provides a linear visualization of the graph (\S\ref{sec:viz}).
\item \textit{odgi draw} renders a 2D image of the graph (\S\ref{sec:viz}).
\item \textit{odgi extract} excerpts subsets of the graph based on path ranges (\S\ref{sec:extract}).
\item \textit{odgi explode} breaks the graph into connected components (\S\ref{sec:extract}).
\item \textit{odgi squeeze} unifies disjoint graphs (\S\ref{sec:extract}).
\item \textit{odgi chop} breaks long nodes into shorter ones (\S\ref{sec:edit}).
\item \textit{odgi unchop} combines unitig nodes (\S\ref{sec:edit}).
\item \textit{odgi break} removes cycles in the graph (\S\ref{sec:edit}).
\item \textit{odgi prune} removes complex regions (\S\ref{sec:edit}).
\item \textit{odgi groom} resolves spurious inverting links (\S\ref{sec:edit}).
\item \textit{odgi position} lifts coordinates between path and graph positions (\S\ref{sec:untangle}).
\item \textit{odgi server} provides a HTTP server to lift coordinates (\S\ref{sec:untangle}).
\item \textit{odgi untangle} deconvolutes paths relative to a reference (\S\ref{sec:untangle}).
\item \textit{odgi tips} finds path end points relative to a reference (\S\ref{sec:untangle}).
\item \textit{odgi sort} orders the graph nodes (\S\ref{sec:sort}).
\item \textit{odgi layout} establishes a 2D layout (\S\ref{sec:sort}).
\item \textit{odgi matrix} derives the pangenome matrix (\S\ref{sec:metrics}).
\item \textit{odgi paths} lists and extracts paths in FASTA (\S\ref{sec:metrics}).
\item \textit{odgi flatten} converts the graph to FASTA and BED (\S\ref{sec:metrics}).
\item \textit{odgi stats} provides numerical properties of the graph (\S\ref{sec:metrics}).
\item \textit{odgi bin} generates a summarized view of the graph (\S\ref{sec:metrics}).
\item \textit{odgi depth} describes node depth over graph and path positions (\S\ref{sec:metrics}).
\item \textit{odgi degree} describes node degree over graph and path positions (\S\ref{sec:metrics}).
\end{itemize}

@@

A universal coordinate system allows us to support several kinds of ``lift-over'' of coordinates between different genomes in the same or different graphs.
For basic liftover operations, \textit{odgi server} provides a basic HTTP server that can translate very efficiently a path position to a pangenomic position within the same graph.
For more complex coordinates conversions, \textit{odgi position} translates graph and path positions between or within graphs, emitting the liftovers in BED format.
Coordinates can be specified in BED format, but users can even specify a GFF/GTF file to project the annotations into the pangenome graph (Fig.~\ref{fig:odgi_viz}\textbf{f-g}).

>>>>

\begin{itemize}
\item \textit{odgi build} constructs the ODGI data model from GFA file (\S\ref{sec:build}).
\item \textit{odgi view} converts the ODGI data model into GFA file (\S\ref{sec:build}).
\item \textit{odgi viz} provides a linear visualization of the graph (\S\ref{sec:viz}).
\item \textit{odgi draw} renders a 2D image of the graph (\S\ref{sec:viz}).
\item \textit{odgi extract} excerpts subsets of the graph based on path ranges (\S\ref{sec:extract}).
\item \textit{odgi explode} breaks the graph into connected components (\S\ref{sec:extract}).
\item \textit{odgi squeeze} unifies disjoint graphs (\S\ref{sec:extract}).
\item \textit{odgi chop} breaks long nodes into shorter ones (\S\ref{sec:edit}).
\item \textit{odgi unchop} combines unitig nodes (\S\ref{sec:edit}).
\item \textit{odgi break} removes cycles in the graph (\S\ref{sec:edit}).
\item \textit{odgi prune} removes complex regions (\S\ref{sec:edit}).
\item \textit{odgi groom} resolves spurious inverting links (\S\ref{sec:edit}).
\item \textit{odgi position} lifts coordinates between path and graph positions (\S\ref{sec:untangle}).%
%\item \textit{odgi server} provides a HTTP server to lift coordinates (\S\ref{sec:untangle}).
\item \textit{odgi untangle} deconvolutes paths relative to a reference (\S\ref{sec:untangle}).
\item \textit{odgi tips} finds path end points relative to a reference (\S\ref{sec:untangle}).
\item \textit{odgi sort} orders the graph nodes (\S\ref{sec:sort}).
\item \textit{odgi layout} establishes a 2D layout (\S\ref{sec:sort}).
\item \textit{odgi matrix} derives the pangenome matrix (\S\ref{sec:metrics}).
\item \textit{odgi paths} lists and extracts paths in FASTA (\S\ref{sec:metrics}).
\item \textit{odgi flatten} converts the graph to FASTA and BED (\S\ref{sec:metrics}).
\item \textit{odgi stats} provides numerical properties of the graph (\S\ref{sec:metrics}).
\item \textit{odgi bin} generates a summarized view of the graph (\S\ref{sec:metrics}).
\item \textit{odgi depth} describes node depth over graph and path positions (\S\ref{sec:metrics}).
\item \textit{odgi degree} describes node degree over graph and path positions (\S\ref{sec:metrics}).
\end{itemize}

@@

A universal coordinate system allows us to support several kinds of ``lift-over'' of coordinates between different genomes in the same or different graphs.
%For basic liftover operations, \textit{odgi server} provides a basic HTTP server that can translate very efficiently a path position to a pangenomic position within the same graph.
\textit{odgi position} translates graph and path positions between or within graphs, emitting the liftovers in BED format.
Coordinates can be specified in BED format, but users can even specify a GFF/GTF file to project the annotations into the pangenome graph (Fig.~\ref{fig:odgi_viz}\textbf{f-g}).
